---
title: "Chapter 4"
bibliography: references.bib
engine: knitr
---
```{r echo=FALSE}
knitr::opts_chunk$set(message = F, warning = F)
rm(list = ls())
```
## Tools required for Chapter 4

## Code Call Out 4.1
**Two-Way Fixed Effects Estimators and Heterogeneous Treatment Effects**
To understand the potential issues related to heterogeneous treatment effects over time and two-way fixed effect estimators, we will examine a pair of numerical examples.  In particular, we will focus on  the composition of the two way FE estimator $\tau$ estimated from:
$$
y_{st} = \gamma_s + \lambda_t + \tau w_{st} + \varepsilon_{st}
$$ {#eq-twfe}
where $y_{st}$ is the outcome variable, $\gamma_s$ and $\lambda_t$ are state (unit) and time fixed effects, $w_{st}$ is the binary treatment variable that takes the value of 1 if a state (unit) $s$ is treated at time $t$ and otherwise takes 0.  We will work with a quite tractable example based on three units and 10 time periods, and will document how the approaches taken by @GoodmanBacon2018 and by @deChaisemartinDhaultfoeuille2019 to understand the two-way FE estimator compare.

The results from @GoodmanBacon2018 and those from @deChaisemartinDhaultfoeuille2019 are similar, however they take quite different paths to get there. Goodman-Bacon's (like that laid out in @AtheyImbens2018) is "mechanical" in that it is based on the underlying difference-in-differences comparisons between all groups. The result in @deChaisemartinDhaultfoeuille2019 is based on a potential outcomes frame-work, and counterfactuals under parallel trend assumptions. Thus to examine how these methods work requires somewhat different frameworks. In the case of @GoodmanBacon2018, we should consider all possible DD comparisons, while in the case of @deChaisemartinDhaultfoeuille2019 we should consider the treatment effect for each unit and time period,  which requires knowing the observed and counterfactual state.  While the approaches the two papers take to understand the content of the estimator differ, they refer to the same estimator, so always recover the same parameter estimate.  To examine this in a more applied way, we will look at a simulated example.

To do this, let's consider a panel of 3 states/areas over the 10 years ($t$) of 2000 to 2009. One of these units is entirely untreated ($unit = 1$ or group $U$), one is treated at an early time period, 2003, ($unit = 2$ or group $k$), and the other is treated at a later time period, 2006, ($unit = 3$ or group $l$).  We will construct a general structure for this data below:

```{r}
Data <- data.frame(unit = ceiling(1:30/10), year = rep(2000:2009, 3))
head(Data)
```

We will consider a simple-case where the actual data-generating process is known as: $$y_{unit,t} = 2 + 0.2 \times (t - 2000) + 1 \times unit + \beta_1 \times post \times unit + \beta_2 \times post \times unit \times (t - treat).$$ Here $unit$ refers to the unit number listed above (1, 2 or 3), $post$ indicates that a unit is receiving treatment in the relevant time period $t$, and $treat$ refers to the treatment period (2003 for unit 2, and 2006 for unit 3).  Let's generate treatment, time to treatment, and post-treatment variables in `R`:

```{r}
Data$treat <- ifelse(Data$unit == 2, 2006, ifelse(Data$unit == 3, 2003, 0))
Data$time  <- ifelse(Data$treat == 0, 0, Data$year - Data$treat)
Data$post  <- ifelse(Data$time >= 0 & Data$treat != 0, 1, 0)
```

This specification allows for each unit to have its own fixed effect, given that $unit$ is multiplied by 1, and allows for a general time trend increasing by 0.2 units each period across the whole sample. These parameters are not so important, as what we care about are the treatment effects themselves.  The impact of treatment comes from the units $\beta_1$ and $\beta_2$. The first of these, $\beta_1$, captures an immediate unit-specific jump when treatment is implemented which remains stable over time. The second of these, $\beta_2$, implies a trend break occurring *only* for the treated units once treatment comes into place. We will consider 2 cases below. In the first case $\beta_1 = 1$ and $\beta_2 = 0$ (a simple case with a constant treatment effect per unit):

```{r}
Data$y1 <- 2 + (Data$year - 2000) * 0.2 + 1 * Data$unit + 1 * Data$post * Data$unit + 
  0 * Data$post * Data$unit * (Data$time)
```

and in a second case $\beta_1 = 1$ and $\beta_2 = 0.45$.  This is a more complex case in which there are heterogeneous treatment effects over time:

```{r}
Data$y2 <- 2 + (Data$year - 2000) * 0.2 + 1 * Data$unit + 1 * Data$post * Data$unit +
  0.45 * Data$post * Data$unit * (Data$time)
```

These two cases are plotted next where the line with empty circles refers to group $U$, the line with black filled circles refers to group $k$ and the line with squares refers to group $l$

```{r fig.width=10}
#| code-fold: true
#| code-summary: "Show the plot code"
library(ggplot2)
library(ggpubr)
p1 <- ggplot(data = Data, aes(x = year, y = y1, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5) +
  geom_point(aes(shape = as.factor(unit)), size = 2) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2),
                     labels = seq(from = 0, to = 12, by = 2),
                     limits = c(0,12)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
t1 <- ggplot() + geom_text(aes(x = 0, y = 0, label = "(a) Simple Decomposition")) +
  theme_void()
p2 <- ggplot(data = Data, aes(x = year, y = y2, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5) +
  geom_point(aes(shape = as.factor(unit)), size = 2) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5),
                     labels = seq(from = 0, to = 20, by = 5),
                     limits = c(0,20)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
t2 <- ggplot() + geom_text(aes(x = 0, y = 0, label = "(b) Decomposition with trends")) +
  theme_void()
ggarrange(plotlist = list(p1, p2, t1, t2), ncol = 2, nrow = 2, heights = c(0.9, 0.1))
```

```{r echo=F}
rm(p1, p2, t1, t2)
```

### The Two-way Fixed Effect Estimator
First we will estimate the parameter by two-way fixed effects regression.  This will provide us with the parameter estimate that both @GoodmanBacon2018 and @deChaisemartinDhaultfoeuille2019 will construct in a piece-wise fashion.  This is done relatively simply in `R`.  We simply estimate @eq-twfe by linear regression using `lm` as laid out below:

XXXXX CODE HERE TO ESTIMATE EQUATION 1 ABOVE FOR BOTH y1 AND y2XXXXX

Here we see that the coefficient of interest is 2.454545.  We can see that this is between the two unit-specific jumps that occur with treatment (2 and 3).  We will see below why it takes this particular weighted average. 


### @GoodmanBacon2018 Decomposition
Using the values simulated above, let's see how the @GoodmanBacon2018 decomposition allows us to understand estimated treatment effects.  We will consider both:  
- (a) Simple Decomposition  
- (b) Decomposition with trends  

The methodology @GoodmanBacon2018 decomposition suggests that we should calculate all $2 \times 2$ combinations of states and time where post-treatment units are compared to "untreated" unit (laid out at more length in the boo).  In this example, this provides four specific effects, which contribute to $\widehat{\tau}$ as a weighted mean. The specific effects desired are:  

- A. $\widehat{\beta}^{2\times2}_{kU}$ from the comparison of the early treated unit with the untreated unit.  
- B. $\widehat{\beta}^{2\times2}_{lU}$, from the comparison of the latter treated unit with the untreated unit.  
- C. $\widehat{\beta}^{2\times2,k}_{kl}$, from the comparison of the early and latter treated units, when the early unit begin to be treated.  
- D. $\widehat{\beta}^{2\times2,l}_{kl}$, from the comparison of the early and latter treated units, when the latter unit begin to be treated.  

These will then be weighted as laid out in @GoodmanBacon2018 to provide the regression-based estimate.  

#### (a) Simple Decomposition

In this case the @GoodmanBacon2018 methodology estimate $\widehat{\tau}$ weighting the next four DD comparisons

```{r}
#| code-fold: true
#| code-summary: "Show the plot code"
library(dplyr)
p1 <- ggplot(data = Data, aes(x = year, y = y1, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(1,0.1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2),
                     labels = seq(from = 0, to = 12, by = 2),
                     limits = c(0,12)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p2 <- ggplot(data = Data, aes(x = year, y = y1, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(1,1,0.1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2),
                     labels = seq(from = 0, to = 12, by = 2),
                     limits = c(0,12)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p3 <- Data %>% filter(year < 2006) %>%
  ggplot(aes(x = year, y = y1, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(0.1,1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2),
                     limits = c(2000,2009)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2),
                     labels = seq(from = 0, to = 12, by = 2),
                     limits = c(0,12)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p4 <- Data %>% filter(year >= 2003) %>%
  ggplot(aes(x = year, y = y1, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(0.1,1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2),
                     limits = c(2000,2009)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2),
                     labels = seq(from = 0, to = 12, by = 2),
                     limits = c(0,12)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
t1 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "A. Early Group v/s Untreated Group"), size = 3) +
  theme_void()
t2 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "B. Later Group v/s Untreated Group"), size = 3) +
  theme_void()
t3 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "C. Early Group v/s Later Group Before 2006"), 
            size = 3) +
  theme_void()
t4 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "D. Early Group v/s Later Group After 2003"), 
            size = 3) +
  theme_void()
ggarrange(plotlist = list(t1, t2, p1, p2, t3, t4, p3, p4), ncol = 2, nrow = 4, 
          heights = c(0.1, 0.4, 0.1, 0.4))
```

```{r echo=FALSE}
rm(p1, p2, p3, p4, t1, t2, t3, t4)
```

As seen in the plots, in the simple decomposition these effects are constants of 3 and 2 for early and later treated units given that the "treatment effect" is simply $1 \times unit$ in each case.  

##### A. Early Group v/s Untreated Group

In order to calculate the effects we start making the simple DD comparison of the untreated group $U$ ($unit = 1$) with the early treated group $k$ ($unit = 3$) getting $\widehat{\beta}^{2 \times 2}_{kU}$ as $$\widehat{\beta}^{2 \times 2}_{kU} = \left( \overline{y}_k^{Post(k)} - \overline{y}_k^{Pre(k)} \right) - \left( \overline{y}_U^{Post(k)} - \overline{y}_U^{Pre(k)} \right)$$ Where $\overline{y}_k^{Post(k)}$ is the mean of the outcome variable for the early treated group $k$ ($unit = 3$) posterior to treatment, from 2003, $\overline{y}_k^{Pre(k)}$ is the mean for of the outcome variable for the early treated group $U$ ($unit = 3$) prior to treatment, (up until 2002), and $\overline{y}_U^{Post(k)}, \overline{y}_U^{Post(k)}$ are the analogous quantities for the untreated group $U$ ($unit = 1$)

```{r}
(mean(Data$y1[Data$unit == 3 & Data$post == 1]) -
   mean(Data$y1[Data$unit == 3 & Data$post == 0])) -
  (mean(Data$y1[Data$unit == 1 & Data$year >= 2003]) -
     mean(Data$y1[Data$unit == 1 & Data$year < 2003]))
```

This result also can be obtained from the linear regression with the canonical DD formula $$y_{unit,t} = \alpha_0 + \alpha_1 \times Post(k) + \alpha_2 \times \mathbf{1}(unit = 3) + \beta_{kU}^{2\times2} \times Post(k) \times \mathbf{1}(unit = 3) + \varepsilon_i$$ Where $Post(k)$ indicates that the year is equal or greater than the year where the group $k$ ($unit = 3$) received the treatment (2003) and $\mathbf{1}(unit = 3)$ indicates if the observation is from the early treated group $k$ ($unit = 3$)

```{r}
summary(lm(y1 ~ factor(year >= 2003) + factor(unit) + factor(year >= 2003):factor(unit), 
           data = Data, subset = (unit != 2)))
```

A third way to obtain this is from the next linear regression $$y_{unit,t} = \alpha_0 + \beta_{kU}^{2 \times 2} \times Post + \sum_{i = 2001}^{2009} \alpha_{i-2000} \times \mathbf{1}(year = i) + \alpha_{10} \times \mathbf{1}(unit = 3) + \varepsilon_i$$ Where in this case $Post$ indicates if the unit is treated (note for group $U$ this will be always 0), $\mathbf{1}(year = i)$ indicates if the observation is in period $i \in \{2001, \ldots, 2009\}$ and $\mathbf{1}(unit = 3)$ keep its meaning

```{r}
summary(lm(y1 ~ post + factor(year) + factor(unit), data = Data, subset = (unit != 2)))
```

Now we store this result for posterior use

```{r}
bku <- lm(y1 ~ post + factor(year) + factor(unit), data = Data,
          subset = (unit != 2))$coefficient["post"]
```

##### B. Later Group v/s Untreated Group
The next DD comparison we calculate is that which compares the later treated group $l$ ($unit = 2$) with the untreated group $U$ ($unit = 1$), resulting in $\widehat{\beta}^{2 \times 2}_{lU}$.  As above, we can generate this DD estimate in a number of ways (most simply by double-differencing with means), and this will then be stored.

```{r}
blu <- lm(y1 ~ post + factor(year) + factor(unit), data = Data, 
   subset = (unit != 3))$coefficient["post"]
blu
(mean(Data$y1[Data$unit == 2 & Data$post == 1]) -
   mean(Data$y1[Data$unit == 2 & Data$post == 0])) -
  (mean(Data$y1[Data$unit == 1 & Data$year >= 2006]) -
     mean(Data$y1[Data$unit == 1 & Data$year < 2006]))
summary(lm(y1 ~ factor(year >= 2006) + factor(unit) + factor(year >= 2006):factor(unit), 
           data = Data, subset = (unit != 3)))
```

##### C. Early Group v/s Later Group Before 2006

Next we calculate the effects from the DD comparisons of early and later treated groups, up until the later treated group receives treatment (2006).  This is:
$$\widehat{\beta}^{2 \times 2, k}_{kl} \equiv \left( \overline{y}^{Mid(k,l)}_{k} - \overline{y}^{Pre(k)}_{k} \right) - \left( \overline{y}^{Mid(k,l)}_{l} - \overline{y}^{Pre(k)}_{l} \right)$$
where $\overline{y}^{Mid(k,l)}_{k}$ is the mean of the outcome variable for the early treated group $k$ ($unit = 3$) in the period between the treatment for the group $k$ and the group $l$ ($unit = 2$), from 2003 to 2005, $\overline{y}^{Pre(k)}_{k}$ is the mean for of the outcome variable for the early treated group $k$ ($unit = 3$) previous to treatment, until 2002, and $\overline{y}^{Mid(k,l)}_{l}, \overline{y}^{Pre(k)}_{l}$ are the analogous for the later treated group $l$ ($unit = 2$)

```{r}
bklk <- lm(y1 ~ post + factor(year) + factor(unit), data = Data,
   subset = (unit != 1 & year < 2006))$coefficient["post"]
bklk
(mean(Data$y1[Data$unit == 3 & (Data$year >= 2003 & Data$year < 2006)]) -
   mean(Data$y1[Data$unit == 3 & Data$year < 2003])) -
  (mean(Data$y1[Data$unit == 2 & (Data$year >= 2003 & Data$year < 2006)]) -
     mean(Data$y1[Data$unit == 2 & Data$year < 2003]))
summary(lm(y1 ~ factor(year >= 2003) + factor(unit) + factor(year >= 2003):factor(unit), 
           data = Data, subset = (unit != 1 & year < 2006)))
```

##### D. Early Group v/s Later Group After 2003

The last DD comparison is for early and later treated groups, starting from 2006 $$\widehat{\beta}^{2 \times 2, l}_{kl} \equiv \left( \overline{y}^{Post(l)}_{l} - \overline{y}^{Mid(k,l)}_{l} \right) - \left( \overline{y}^{Post(l)}_{k} - \overline{y}^{Mid(k,l)}_{k} \right)$$ Where $\overline{y}^{Post(l)}_{l}$ is the mean of the outcome variable for the later treated group $l$ ($unit = 2$) in the period after this group received the treatment, from 2006, $\overline{y}^{Mid(k,l)}_{l}$ is the mean for of the outcome variable for the later treated group $l$ ($unit = 2$) in the period between the treatment for the group $k$ ($unit = 3$) and the group $l$, from 2003 to 2005, and $\overline{y}^{Post(l)}_{k}, \overline{y}^{Mid(k,l)}_{k}$ are the analogous quantities for the early treated group $k$ ($unit = 3$).  We can generate and save this quantity as we have previously:

```{r}
bkll <- lm(y1 ~ post + factor(year) + factor(unit), data = Data,
   subset = (unit != 1 & year > 2002))$coefficient["post"]
bkll
(mean(Data$y1[Data$unit == 2 & Data$year > 2005]) -
   mean(Data$y1[Data$unit == 2 & (Data$year >= 2003 & Data$year < 2006)])) -
  (mean(Data$y1[Data$unit == 3 & Data$year > 2005]) -
     mean(Data$y1[Data$unit == 3 & (Data$year >= 2003 & Data$year < 2006)]))
summary(lm(y1 ~ factor(year >= 2006) + factor(unit) + factor(year >= 2006):factor(unit==2), 
           data = Data, subset = (unit != 1 & year > 2002)))
```

This comparison is the comparison which can potentially result in undesired results if treatment effects are dynamic over time because it views group 3 (the previously treated group) as a control.  However, in this case, given that treatment effects are homogenous over time we do not have a major problem here, and we observe that $\widehat{\beta}^{2 \times 2, l}_{kl}=2$.

##### Weights

We can now arrive to the OLS estimate of this two-way fixed effect model by generating the weighted mean of the previous estimates as: $$\widehat{\tau} = W_{kU} \cdot \widehat{\beta}^{2\times 2}_{kU} + W_{lU} \cdot \widehat{\beta}^{2\times 2}_{lU} + W_{kl}^{k} \cdot \widehat{\beta}^{2\times 2,k}_{kl} + W_{kl}^{l} \cdot \widehat{\beta}^{2\times 2,l}_{kl}$$ Where each $W$ is the weight that the respective $\beta$ has in this weighted mean, specifically:
\begin{align*} 
W_{kU} & = \frac{(n_k + n_U)^2\widehat{V}^D_{kU}}{\widehat{V}^D} \quad &  \quad W_{lU} & = \frac{(n_l + n_U)^2\widehat{V}^D_{lU}}{\widehat{V}^D} \\ 
W_{kl}^k & = \frac{[(n_k + n_l)(1 - \overline{D}_l)]^2\widehat{V}^{D,k}_{kl}}{\widehat{V}^D} \quad &  \quad W_{kl}^l & = \frac{[(n_k + n_l)(1 - \overline{D}_k)]^2\widehat{V}^{D,l}_{kl}}{\widehat{V}^D}
\end{align*}
Where $n$ refers to the sample share of the group

```{r}
nk = 1/3
nl = 1/3
nu = 1/3
```

$\overline{D}$ referes to the share of time the group is treated

```{r}
Dk = mean(Data$post[Data$unit==3])
Dl = mean(Data$post[Data$unit==2])
```

and $\widehat{V}$ refers to how much treatment varies

```{r}
VkU = 0.5*0.5*(Dk)*(1-Dk)
VlU = 0.5*0.5*(Dl)*(1-Dl) 
Vklk = 0.5*0.5*((Dk-Dl)/(1-Dl))*((1-Dk)/(1-Dl))
Vkll = 0.5*0.5*(Dl/Dk)*((Dk-Dl)/(Dk))
VD   = sum(lm(post ~ factor(unit) + factor(year), 
              data = Data)$residuals^2)/30
```

The weights are thus the following:

```{r}
wkU = ((nk + nu)^2*VkU)/VD
wkU
wlU = ((nl + nu)^2*VlU)/VD
wlU
wklk = (((nk + nl)*(1-Dl))^2*Vklk)/VD
wklk
wkll = (((nk + nl)*Dk)^2*Vkll)/VD
wkll
```

With this in mind the $\tau$ estimate is 

```{r}
tau = wkU * bku + wlU * blu + wklk * bklk + wkll * bkll
tau
```
as observed in the two-way fixed effect estimate above.

#### (b) Decomposition with trends

In this case the @GoodmanBacon2018 decomposition follows as above generating the treatment effect as follows:

```{r}
#| code-fold: true
#| code-summary: "Show the plot code"
library(dplyr)
p1 <- ggplot(data = Data, aes(x = year, y = y2, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(1,0.1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5),
                     labels = seq(from = 0, to = 20, by = 5),
                     limits = c(0,20)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p2 <- ggplot(data = Data, aes(x = year, y = y2, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(1,1,0.1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2)) +
  scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5),
                     labels = seq(from = 0, to = 20, by = 5),
                     limits = c(0,20)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p3 <- Data %>% filter(year < 2006) %>%
  ggplot(aes(x = year, y = y2, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(0.1,1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2),
                     limits = c(2000,2009)) +
  scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5),
                     labels = seq(from = 0, to = 20, by = 5),
                     limits = c(0,20)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
p4 <- Data %>% filter(year >= 2003) %>%
  ggplot(aes(x = year, y = y2, color = as.factor(unit))) +
  geom_line(linetype = 1, size = 0.5, aes(alpha = as.factor(unit))) +
  geom_point(aes(shape = as.factor(unit), alpha = as.factor(unit)), size = 2) +
  scale_alpha_manual(values = c(0.1,1,1)) +
  scale_shape_manual(values = c(1, 16, 12)) +
  scale_color_manual(values = c("black", "black", "black")) +
  labs(x = "Time", y = "Outcome Variable") +
  scale_x_continuous(breaks = seq(from = 2000, to = 2009, by = 2),
                     limits = c(2000,2009)) +
  scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5),
                     labels = seq(from = 0, to = 20, by = 5),
                     limits = c(0,20)) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_vline(xintercept = 2005, color = "red", linetype = 2) +
  theme(legend.position = "none")
t1 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "A. Early Group v/s Untreated Group"), size = 3) +
  theme_void()
t2 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "B. Later Group v/s Untreated Group"), size = 3) +
  theme_void()
t3 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "C. Early Group v/s Later Group Before 2006"), 
            size = 3) +
  theme_void()
t4 <- ggplot() + 
  geom_text(aes(x = 0, y = 0, label = "D. Early Group v/s Later Group After 2003"), 
            size = 3) +
  theme_void()
ggarrange(plotlist = list(t1, t2, p1, p2, t3, t4, p3, p4), ncol = 2, nrow = 4, 
          heights = c(0.1, 0.4, 0.1, 0.4))
```

```{r echo=FALSE}
rm(p1, p2, p3, p4, t1, t2, t3, t4)
```

As seen in the plots, in the decomposition with trends these effects are no longer constants of 3 and 2 for early and later treated units given that the "treatment effect" is no longer simply $1 \times unit$ in each case.  

```{r}
# 2X2 DD Regressions
A <- lm(y2 ~ post + factor(year) + factor(unit), data = Data, subset=(unit!=2))
B <- lm(y2 ~ post + factor(year) + factor(unit), data = Data, subset=(unit!=3))
C <- lm(y2 ~ post + factor(year) + factor(unit), data = Data, subset=(unit!=1 & year<2006))
D <- lm(y2 ~ post + factor(year) + factor(unit), data = Data, subset=(unit!=1 & year>2002))
# 2x2 Betas
bkUk <- A$coefficient["post"]
bkUl <- B$coefficient["post"]
bklk <- C$coefficient["post"]
bkll <- D$coefficient["post"]
bkll
# Share of time treated
Dk = mean(Data$post[Data$unit==3])
Dl = mean(Data$post[Data$unit==2])
# How much treatment varies
VkUk = 0.5*0.5*(Dk)*(1-Dk)
VkUl = 0.5*0.5*(Dl)*(1-Dl) 
Vklk = 0.5*0.5*((Dk-Dl)/(1-Dl))*((1-Dk)/(1-Dl))
Vkll = 0.5*0.5*(Dl/Dk)*((Dk-Dl)/(Dk))
VD <- sum(lm(post ~ factor(unit) + factor(year), data = Data)$residuals^2/30)
# Share of sample
nk   = 1/3
nl   = 1/3
nu   = 1/3
# Weights
wkUk = ((nk + nu)^2*VkUk)/VD
wkUl = ((nl + nu)^2*VkUl)/VD
wklk = (((nk + nl)*(1-Dl))^2*Vklk)/VD
wkll = (((nk + nl)*Dk)^2*Vkll)/VD
# Tau
tau = bkUk*wkUk + bkUl*wkUl + bklk*wklk + bkll*wkll
tau
```
What is noteworthy here is the surprising behaviour flagged by @GoodmanBacon2018 for the final comparison based on the case where the earlier treated unit (unit 3) is used as a control for the later trated unit (unit 2).  In this case, given that there *are* time-varying treatment effects, despite the fact that each unit-specific treatment effect is positive, we observe that the parameter $\widehat{\beta}^{2 \times 2, l}_{kl}$ is actually *negative*.  In this particular example this negative value (-1.375) is not sufficient to turn the weighted treatment effect estimate negative, but if you play around with the size of the parameters $\beta_1$ and $\beta_2$ above, you will see that large enough differences in trends *can* result in such estimates!  Here, as above, we see that when we aggregate unit-specific estimates as `tau`, the estimate (by definition) agrees with the estimate generated by two-way fixed effect models previously.


### @deChaisemartinDhaultfoeuille2019's Procedure
Now, we will show that the procedures described in @deChaisemartinDhaultfoeuille2019, despite arriving to the estimator in a different way, also let us understand how the regression weights the two-way fixed effect estimator.  
